eda

### 2024-05-14 analysis/eda.Rmd
OA analysis Jan 2024

>>>>

## Compare to HOAD

```{r}
wos_jn_all <- wos_oa |>
  group_by(earliest_year) |>
  summarise(journals = n_distinct(issn_l),
            articles = sum(wos_articles))

wos_oa_per_year <-  wos_oa |>
  filter(is_oa == TRUE) |>
  group_by(earliest_year) |>
  summarise(oa_journals = n_distinct(issn_l),
            oa_articles = sum(wos_articles))

wos_per_year <- inner_join(wos_jn_all, wos_oa_per_year)

cr_all_per_year <- hoaddata::jn_ind |> 
  distinct(issn_l, cr_year, jn_all) |>
  group_by(cr_year) |>
  summarise(cr_journals = n_distinct(issn_l),
            cr_articles = sum(jn_all))

cr_oa <- hoaddata::jn_ind |>
  filter(!is.na(cc)) |> group_by(cr_year) |>
  summarise(cr_oa = sum(cc_total))

cr_per_year <- inner_join(cr_all_per_year, cr_oa) |>
  mutate(cr_year = as.numeric(as.character(cr_year)))

my_df_year <- inner_join(cr_per_year, wos_per_year, by = c("cr_year" = "earliest_year")) |>
  mutate(cr_oa_prop = cr_oa / cr_articles,
         wos_oa_prop = oa_articles / articles)

### journal level

wos_jn_all <- wos_oa |>
  group_by(earliest_year, issn_l) |>
  summarise(articles = sum(wos_articles))

wos_oa_per_year <-  wos_oa |>
  filter(is_oa == TRUE) |>
  group_by(earliest_year, issn_l) |>
  summarise(oa_articles = sum(wos_articles))

wos_per_year <- inner_join(wos_jn_all, wos_oa_per_year)

cr_all_per_year <- hoaddata::jn_ind |> 
  distinct(issn_l, cr_year, jn_all) |>
  group_by(cr_year, issn_l) |>
  summarise(cr_articles = sum(jn_all))

cr_oa <- hoaddata::jn_ind |>
  filter(!is.na(cc)) |> group_by(cr_year, issn_l) |>
  summarise(cr_oa = sum(cc_total))

cr_per_year <- inner_join(cr_all_per_year, cr_oa) |>
  mutate(cr_year = as.numeric(as.character(cr_year)))

my_df <- inner_join(cr_per_year, wos_per_year, by = c("cr_year" = "earliest_year", "issn_l")) |>
  mutate(cr_oa_prop = cr_oa / cr_articles,
         wos_oa_prop = oa_articles / articles)
```

### wos_jct_items_approach

use items from kb, distinguishing between doc types and provides oa information

```{r}
wos_jct_items_raw <- readr::read_csv(here::here("data-raw/", "wos_jct_items_df.csv")) 
wos_jct_items <- wos_jct_items_raw|>
  mutate(core_pub = ifelse(grepl("Article|Review", item_type), 1, 0),
         hybrid_oa = ifelse(grepl("hybrid", oa_status), 1, 0),
         cr_oa = ifelse(tolower(doi) %in% hoaddata::cc_articles$doi, 1, 0),
         online_year = lubridate::year(wos_pubdate_online)) |>
  mutate(earlies_year = ifelse(!is.na(online_year), online_year, pubyear)) |>
  distinct()
```

## stats

```{r}
core_hybrid_oa <- wos_jct_items |>
  filter(core_pub == 1) |>
  group_by(earlies_year) |>
  summarise(core_hybrid_oa = sum(hybrid_oa),
            core_cr_oa = sum(cr_oa))

wos_oct_23_stat <- wos_jct_items |>
  group_by(earlies_year) |>
  summarise(all = n_distinct(item_id),
            core = sum(core_pub),
            hybrid_oa = sum(hybrid_oa),
            cr_oa = sum(cr_oa)) |>
  inner_join(core_hybrid_oa) |>
  mutate(oa_prop = hybrid_oa / all,
         core_oa_prop = core_hybrid_oa / core,
         cr_prop = cr_oa / all,
         core_cr_prop = core_cr_oa / core)

write_csv(wos_oct_23_stat, here::here("data", "wos_oct_23_stat.csv"))
```
